Natural variations in water discharge quantity have an inconsistent impact on sediment transport capacity in subglacial channels

### 2023-10-09 Delaneyetal2023b.tex
General changes and comments

\title{Natural variations in water discharge quantity have an inconsistent impact on sediment transport capacity in subglacial channels}

@@

\begin{keypoints}
\item Water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size.
\item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
\item Water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge, as opposed to subglacial sediment transport capacity in glacierized catchments.
\end{keypoints}

\begin{abstract}  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the shear stress exerted on the channel bottom.
  In subaerial channels, width, and velocity evolve quickly with water discharge, while in subglacial channels, water discharge changes mainly affect water velocity due to slow-evolving channel geometry.
  Parameterizations of these subglacial and subaerial hydraulics are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet revealing greater variability in sediment transport capacity in subglacial channels.
  Furthermore, results show that high subglacial sediment transport rates can persist across various water discharges.
  Slowly evolving channel width accommodates some variability in subglacial channels.
  Despite differences in response to water discharge, sediment discharge measurements generally correlate better with subaerial hydraulic parameters.
  This implies that other factors like sediment access or proglacial processes might be at play, affecting sediment transport capacity differently in subglacial and subaerial channels.
\end{abstract}

@@

In a transport-limited regime, sediment discharge responds to the sediment transport capacity, defined as the amount of sediment the water can carry.
Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which to mobilize sediment.
The shear stress  across the channel width $\tau_t$ responds to the velocity of water $v$ flowing through the channel so that
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau_t}
    \tau_t \propto w\, v^2,
  \end{equation}
\end{linenomath*}
%
where $w$ is the width of the channel. Following mass conservation, the velocity of the water flowing through a  channel is
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q}{S},
  \end{equation}
\end{linenomath*}
where $Q$ is water discharge,  and $S$ is the channel's wetted area.

In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \cite{leopold1953}.
The change in water depth results in a commensurate increase in water velocity.
Thus,  fulfilling Equation~\ref {eq:v}, the higher discharge is accommodated by both an increase in wetted area and flow velocity.
As a result, the shear stress across the channel $\tau_t$ increases according to Equation~\ref{eq:tau_t}.

The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by frictional heating of water flowing through the channel \cite{rothlisberger1972}.
As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
Therefore, subglacial water flow behaves more like pipe flow, and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w$ evolves as well.

  \begin{figure}[h]
  \centering
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different responses that subglacial and subaerial channels to increased water discharge.
      Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
      $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
      The subglacial channel widths $w_g$ remain unchanged, while the subaerial channel widths $w_f$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
      %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
    }
    \label{fig:cartoon}
  \end{figure}
% \ian{Come back to Daniel's comments}

As a result, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
These differences are implicitly included in a wide range of available models that  quantify sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016}.
Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels \cite{alley1997}.
%Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins. %\mauro{self-cite here?}

This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels and compares it to subaerial ones.
In the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized.
Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
The parameterizations' lumped nature isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
Furthermore, outputs are compared to sediment discharge data from these two catchments.
The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport from glacierized catchments.

@@

To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel parameterization takes into account the channel geometry and the velocity of the flowing water.
To do this, a modified version of the lumped hydraulics model presented in \citeA{clarke1996} and \citeA{werder2010} is used.

Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to frictional heating from water flow and creep closure by ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
The formulation here does not consider the englacial storage of water.
Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution.
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of the subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.

Following the Darcy-Weisbach equation, head drop $\Delta h$ is,
\begin{linenomath*}
  \begin{equation}
    \label{eq:dh}
    \Delta h \,  = l \,\frac{1}{g}\,s\,f_r\,\frac{Q^2}{D_h^5}
  \end{equation}
\end{linenomath*}
\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, and $l$ is the length over which head change occurs. $s$ is a factor accounting for channel geometry \cite{hooke1990}, calculated as
\begin{equation}
  \label{eq:Hf}
  s = \frac{2\,(\beta -\sin \beta)^2}{(\frac{\beta}{2}\,+\,\sin \frac{\beta}{2})^4},
\end{equation}
where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle). Note that $\beta =\pi$ corresponds to a semi-circle and
smaller values of $\beta$ result in shallow, wide channels.
The hydraulic diameter $D_h$ is converted to wetted area $S_g$ given a channel geometry following \cite{hooke1990}
\begin{equation}
  \label{eq:Dh2S}
  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta}.
\end{equation}

With wetted area $S_g$, the shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
\begin{equation}
  \label{eq:tau}
  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
\end{equation}
%
where $v_g = \frac{Q}{S_g}$ is the water velocity.

The width-integrated shear stress is represented as
\begin{equation}
  \label{eq:tautg}
  \tau_{tg}=w_g\,\tau_g,
\end{equation}

where  channel floor width $w_g$ is
\begin{equation}
  \label{eq:dh2wc}
  w_g = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
\end{equation}

@@

To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
the shear stress $\tau_f$ at the river bed is
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels.
Channel width $w_f$ is
\begin{equation}
  \label{eq:wcf}
  w_f = k \, Q^{\alpha},
\end{equation}
%
where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
%
As above, the width-integrated shear stress is
\begin{equation}
    \label{eq:tautf}
    \tau_{tf}=w_f\,\tau_f.
  \end{equation}

@@

The parameterizations above are applied to hydrological records from the Fieschergletscher (\alpine) and the Leverett Glacier (\icesheet).
The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
Outputs of the parameterizations represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
To generalize these cases, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
\icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).

For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{2}$.
Friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable water velocities for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
In both subaerial channels, the bed slope is $\nabla z_c = 0.05$.
% Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.

Second, to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels, $100$  parameterizations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$).
Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies between $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies between $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013}.
The variability of output is represented by the  standard deviation of that output over the course of the season.

Lastly, outputs are compared with the sediment discharge records of the glaciers using Spearman rank correlation, dependent on the ordering of values.
Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing.
Outputs and observations at $15$\,\unit{min} intervals are aggregated across timescales of up to $15$\,\unit{days} by averaging values one-half of the timescale before and after a point time.

@@

\subsection{Role of water discharge in sediment transport capacity}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
  \caption{Parameterization outputs from the hydrographs n panels a and e for \alpine cases (a-d) and \icesheet{} (e-h).
    Gray (black) lines in a and e represent the subglacial channel size (water discharge).
    Blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
    Data are shown at $15$\unit{min} intervals.
    Arrows denote where variables peaking in the subglacial channel prior to the subaerial one.
    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area.
    %Note that the x and y axes are different for \alpine{} and \icesheet{}.
  }
  \label{fig:model_outs}
\end{figure}

Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

In the subglacial channel, peaks generally occur when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}).
Alternatively, in the subaerial channel,  sediment transport variables peak  hen water discharge is highest.
The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge.
Furthermore, the subglacial outputs are generally larger than the subaerial values, thus more efficient at transporting sediment.
Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig3.png}
  \caption{
    Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
    These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
    Variables have been normalized to mean values.
    $R_g$ and $R_f$ show the Spearman rank correlations for the subglacial and the subaerial outputs, respectively.
    Plots are shown with data and outputs at $15$\unit{min} intervals.
  }
  \label{fig:Qw_vari}
\end{figure}

Each water discharge value in the subaerial channel results in a unique water velocity, shear stress, and width-integrated shear stress, creating a perfect rank correlation between the variables (Figure~\ref{fig:Qw_vari} a--f).

In the subglacial channel, however, the persistence of high water velocity, shear stress, and width-integrated shear stress across a range of discharges suggests that substantial sediment transport may occur across  (Figure~\ref{fig:Qw_vari} a--f).

>>>>

\title{Natural variations in water discharge quantity have an inconsistent impact on sediment transport capacity in subglacial channels\mauro{\\Title does not work.  I don't think ``inconsistent'' can be used like this. Maybe:}\\
Subglacial and subaerial fluvial sediment transport respond differently to discharge variations
}

@@

\begin{keypoints}
\item Short term water discharge variations in subglacial channels are mainly accommodated by water velocity, not channel size.
\item Greater variability in water velocity causes sediment transport capacity to vary more in subglacial channels than in subaerial channels.
\item Water discharge may represent processes including proglacial sediment mobilization and sediment access driving variations in sediment discharge, as opposed to subglacial sediment transport capacity in glacierized catchments. \mauro{I don't understand this sentence.}
\end{keypoints}

\begin{abstract}  Sediment transport capacity in subglacial and subaerial channels depends on  water velocity and channel width, which together control the total shear stress exerted on the channel bottom.
  In subaerial channels water discharge variations are accommodated largely by changes in flow depth and width.
  While in subglacial channels, water discharge changes primarily affect water velocity due to slow-evolving channel geometry.
  Parameterizations of these subglacial and subaerial hydraulics are applied to hydrographs from an Alpine glacier and the Greenland Ice Sheet revealing greater variability in the inferred sediment transport capacity in subglacial than in subaerial channels.
  Results show that high subglacial sediment transport rates can occur across a wide range of water discharges.
%  Slowly evolving channel width accommodates some variability in subglacial channels. \mauro{not sure what that sentence should do}
  Despite differences in response to water discharge, sediment discharge measurements generally correlate better with subaerial hydraulic parameters.
  This implies that other factors like sediment access or proglacial processes might be at play that affect sediment transport capacity differently in subglacial and subaerial channels.
\end{abstract}

@@

In a transport-limited regime, sediment discharge tracks sediment transport capacity, defined as the amount of sediment the water can carry.
Sediment transport capacity depends on the shear stress between water and the sediment it flows  over \cite{shields1936,meyer1948,engelund1967} along with the width of the channel bottom $w$ over which to mobilize sediment.
The shear stress $\tau$ responds to the velocity of water $v$ flowing through the channel so that
\begin{linenomath*}
  \begin{equation}
    \label{eq:tau}
    \tau_t \propto v^2.
  \end{equation}
\end{linenomath*}
\mauro{I removed the width from here as I found it odd to introduce that here.}
%
Following mass conservation, the velocity of the water flowing through a  channel is
\begin{linenomath*}
  \begin{equation}
    \label{eq:v}
    v = \frac{Q}{S},
  \end{equation}
\end{linenomath*}
where $Q$ is water discharge,  and $S$ is the channel's wetted area. \mauro{I find it odd to have formulas in the Intro}

In subaerial channels, operating with open channel flow, the wetted area $S$ of the channel evolves with changing water discharge $Q$, by changing both the channel's width and the water depth \cite{leopold1953}.
The change in water depth results in a commensurate increase in water velocity.
%Thus,  fulfilling Equation~\ref {eq:v}, the higher discharge is accommodated by both an increase in wetted area and flow velocity. \mauro{not needed}
As a result, the shear stress $\tau$ increases according to Equation~\ref{eq:tau}.

The response of water velocity to changing water discharge in subglacial channels differs from subaerial ones.
The size of subglacial channels responds to the creep closure of the ice above and the opening of the channel by melting due to frictional heating of water flowing through the channel \cite{rothlisberger1972}.
As a result, the subglacial channel size only evolves relatively slowly over days to months, whereas water discharge can vary over hours \cite<e.g.>{iken1986,andrews2014,nanni2020}.
Therefore, on short time scales, subglacial water flow behaves more like pipe flow and changes in water discharge $Q$ are mainly accommodated with increased or decreased water velocity $v$ \cite{alley1997} because the size of the channel $S$ changes on a much longer timescale (Equation~\ref{eq:v} and Figure~\ref{fig:cartoon}).
%It follows that an increase in water discharge $Q$ will cause the shear stress across the channel $\tau_t$ to respond with water velocity $v$, as opposed to subaerial channels where channel width $w$ evolves as well.

  \begin{figure}[h]
  \centering
    \includegraphics[width=0.8\linewidth]{Fig1.pdf}
    \caption{Sketch for the different responses that subglacial and subaerial channels to increased water discharge.
      Water velocity magnitudes in the subglacial, $v_g$, and subaerial, $v_f$, channels are shown by arrow length.
      $S_g$ and $S_f$ represent the wetted area in subglacial and subaerial channels, respectively.
      The subglacial channel widths $w_g$ remain unchanged, while the subaerial channel widths $w_f$ evolve. Shear stress $\tau$ is responsible for the mobilization of sediment.
      %Note that in the subaerial channel parameterization, channels have rectangular shapes with their width much larger than their depth (Section~\ref{sect:fluv}).
\mauro{Needs renaming $w_{cg/f}$ to $w_{g/f}$. Then, I find the downward arrow confusing.  Maybe just leave away? Also, I would write $\tau_f$ and $\tau_g$, it's ok to not have the width integrated tau here (and maybe we remove $\tau_{tf}$ from the paper anyway...}
    }
    \label{fig:cartoon}
  \end{figure}
% \ian{Come back to Daniel's comments}

Consequently, sediment mobilization in subaerial and subglacial channels responds differently to changing water discharge.
These differences are implicitly included in a wide range of available models that  quantify sediment transport in both subglacial and subaerial channels  \cite<e.g.>{walder1994,alley1997,tucker1997,creyts2013,delaney2019,hewitt2019,wickert2019}.
The divergent response of changing water discharge on sediment mobilization capacity may well impact sediment dynamics at glacier margins where flow transitions from pressurized to open channel flow \cite<e.g.>{perolo2018}.
Additionally, the variable response of sediment transport capacity to water discharge in the two systems may affect the interpretation of sediment transport records in glacial systems \cite<e.g.>{muller1968,richards2003,swift2005,ganti2016}.
Yet, little work has examined the differing relationship between variations in water discharge and sediment discharge in subglacial and subaerial channels \cite{alley1997}.
%Furthermore, thorough observations of water velocity and channel width in subglacial channels are limited and, to these authors' knowledge, not contrasted with subaerial conditions at the glacier margins. %\mauro{self-cite here?}

This manuscript evaluates the relationship between sediment transport capacity and water discharge in subglacial channels and compares it to subaerial ones.
Due to the absence of continuous measurements of subglacial and subaerial water velocity and channel morphology, numerical parameterizations of subglacial and subaerial hydraulics are utilized. \mauro{Aren't those things always parameterised, irrespective of whether there are measurements or not?}
Parameterizations are applied to hydrological records from an Alpine glacier in Switzerland (Fieschergletscher) and  a land-terminating glacier in Greenland (Leverett Glacier).
Our parameterizations' lumped nature isolates the disparity between water discharge and sediment transport capacity in subglacial systems, independent of the upstream drainage network and sediment access.
Outputs demonstrate differences in the relationships between  water discharge, channel geometry, and water velocity in the subglacial and subaerial channels.
Furthermore, outputs are compared to sediment discharge data from these two catchments.
The manuscript then uses the demonstrated differences between the two channel types to discuss the implications for interpreting the processes driving variability in records of sediment transport from glacierized catchments.
\mauro{As stated in the general comments.  I'd rather use ``model'' than ``parameterization'' as the latter is, for me, something which is used in a model and not a model itself.}

@@

To evaluate the shear stress of water flowing across sediments underneath a glacier, the subglacial channel parameterization takes into account the channel geometry and the velocity of the flowing water.
To do this, a lumped hydraulics model \cite{clarke1996} modified from the one presented by \citeA{werder2010b} is used.

Here, it is assumed that the water is transported through a subglacial channel \cite<Figure~\ref{fig:cartoon}; >{rothlisberger1972}, and that the channel  size responds to melt due to frictional heating from water flow and to creep closure of the ice, using the Darcy-Weisbach formulation for water-flow through a pipe  \cite<e.g.>{rothlisberger1972,clarke2003}.
The formulation here does not consider the englacial storage of water.
Thus, it is assumed that the changing hydraulic head at the top of the glacier is negligible in the channel size evolution \mauro{I don't understand this sentence}.
The evolution of subglacial channel size $S_g$ is given as
\begin{linenomath*}
  \begin{equation}
    \label{eq:dS_dt}
    \frac{\partial S_g}{\partial t} = C_1 \frac{Q \Delta h}{l} - C_2 \left(h_{o}-\frac{\Delta h}{2}\right)^n\,S_g,
  \end{equation}
\end{linenomath*}
\noindent where $C_1= (1-\rho_wc_pc_t)\,\frac{\rho_wg}{\rho_iL}$ and $C_2=2A(\frac{\rho_wg}{n})^n$ are constants (Table \ref{table:vpm}), $l$ is the length of the subglacial channel, $Q$ is water discharge, $\Delta h$ is the hydraulic potential change from the glacier terminus to its top, $h_{o}= \frac{\rho_i}{\rho_w} h_{ice}$ is the ice overburden pressure expressed in meter water equivalent, and $n$ is Glen's n \cite{glen1955} (usually $n=3$).
The first term on the right side of the equation represents the opening of the channel through frictional heating, while the following term represents the creep closure of the channel from ice deformation.

Following the Darcy-Weisbach equation, the head drop $\Delta h$ is
\begin{linenomath*}
  \begin{equation}
    \label{eq:dh}
    \Delta h \,  = l \,\frac{1}{2g} \,f_r\,\frac{v_g^2}{D_h},
  \end{equation}
\end{linenomath*}
\noindent where $f_r$ is a friction factor, $D_h$ is the hydraulic diameter, $l$ is the length over which head change occurs, and $v_g=Q/S_g$ is the water flow speed.
%
The hydraulic diameter $D_h$ is converted to wetted area $S_g$ with
\begin{equation}
  \label{eq:Dh2S}
  S_g= \frac{D_h^2}{2}\,\frac{(\frac{\beta}{2}+\sin \frac{\beta}{2})^2}{\beta - \sin \beta},
\end{equation}
where $\beta$ is the central angle of the circular segment that comprises the channel (the so-called Hooke angle, \citeA{hooke1990}). Note that $\beta =\pi$ corresponds to a semi-circular channel and smaller values of $\beta$ result in shallow, wide channels.
%
The shear stress, $\tau_g$, between the water and the channel bed is determined through the Darcy-Weisbach formulation
\begin{equation}
  \label{eq:tau}
  \tau_g=\frac{1}{8}\,f_r\,\rho_w\,v_g^2,
\end{equation}
%
where $v_g = \frac{Q}{S_g}$ is the water velocity.
%
The width-integrated shear stress is represented as
\begin{equation}
  \label{eq:tautg}
  \tau_{tg}=w_g\,\tau_g,
\end{equation}
where  channel floor width $w_g$ is
\begin{equation}
  \label{eq:dh2wc}
  w_g = 2  \sin \frac{\beta}{2} \sqrt{\frac{2\, S_g}{\beta -\sin \beta}}.
\end{equation}
\mauro{If we get rid of $\tau_{tf}$, above needs a bit of massaging.  I'd then move the $w_g$ eq to just after the $S_g$ eq.}

@@

% To parameterize the shear stress of water flowing across sediments in the subaerial channel, the hydraulics parameterization presented in \citeA{tucker1997} is implemented.
Using mass conservation and the Darcy-Weisbach relationship as well as assuming that the channel is wide so that the hydraulic radius is well approximated by the flow depth, then
the shear stress $\tau_f$ at the river bed is
\begin{linenomath*}
  \begin{equation}
    \label{eq:DW_tau}
    \tau_f=\frac{\rho_w\,g^{\frac{2}{3}}\,f_p^{\frac{1}{3}}}{2}\, \Big(\frac{Q}{w_f} \Big)^{\frac{2}{3}} \,\nabla z_c^{\frac{2}{3}},
  \end{equation}
\end{linenomath*}
where $\nabla z_c$ is the channel slope, and $f_p$ is the friction factor for subaerial channels \cite{tucker1997}.
Channel width $w_f$ is
\begin{equation}
  \label{eq:wcf}
  w_f = k \, Q^{\alpha},
\end{equation}
%
where $k$ is a constant and $\alpha=\frac{1}{2}$ is a commonly chosen exponent \cite{leopold1953}.
Water velocity, $v_f$, is given by rearranging Equation \ref{eq:tau} as
\begin{equation}
  \label{eq:vf}
  v_f = \sqrt{\frac{8\,\tau_f}{f_p\,\rho_w}}.
\end{equation}
%
As above, the width-integrated shear stress is
\begin{equation}
    \label{eq:tautf}
    \tau_{tf}=w_f\,\tau_f.
  \end{equation}

  Note that the subaerial channel model \mauro{or ``parameterization''} is a purely algebraic model, whereas the subglacial model comprises a differential equation for the evolution of $S_g$ and thus will have a history dependence on the discharge.

@@

\mauro{I would call the ``cases'' ``scenarios'' instead.  The word ``scenario'' does not appear in the manuscript so far, so it would be unambiguous.  Also, ``cases'' has the disadvantage that ``In both cases,'' is a saying, and thus it is confusing.  I implemented this change.}

The parameterizations above are applied to hydrological records from the Fieschergletscher (scenario \alpine) and the Leverett Glacier (scenario \icesheet).
The subglacial and subaerial parameterizations represent the subglacial flow of water exiting a glacier and transitioning to subaerial flow.
Outputs of the parameterizations represent generalizable sediment transport characteristics from these hydrographs, rather than actual hydraulic conditions.
To generalize these scenarios, \alpine{}  is exemplified by relatively thin ice thickness ($h_{ice}$: $225$\,\unit{m}), low water discharge ($\sim\,10$\,\unit{m}$^3$\,\unit{s}$^{-1}$) and high diurnal variability in water discharge (Figure~\ref{fig:model_outs}, a).
\icesheet{}  is exemplified by thick ice  ($h_{ice}$: $700$\,\unit{m}), high water discharge ($\sim\,300$\,\unit{m}$^3$\,\unit{s}$^{-1}$)  and low diurnal variability in water discharge (Figure~\ref{fig:model_outs}, e).
%
However, both scenarios share a discharge with a similar, relative seasonal amplitude.

For each glacier, the parameterizations are first applied to a reference test case for each glacier that assumes  a subglacial channel with $\beta=\frac{\pi}{2}$, i.e. a segment of a quarter-circle.
Friction factors $f_r$ and $f_p$ are tuned so that the parameterizations reproduce reasonable water velocities for both \alpine and \icesheet{}\cite<$\sim\,1.6\,$\unit{m}$^3$\,\unit{s}$^{-1}$>{werder2010b,chandler2013}.
\mauro{Those $f$ reported in the table are super large!  Maybe an issue down the line...}
The subaerial channels of both scenarios have a bed slope of $\nabla z_c = 0.05$ \mauro{Why do you report a value here but most are hidden in the table?  Maybe just ref the table?}.
% Water velocity, shear stress, and width-integrated shear stress in the subglacial and subaerial channels from this formulation are presented below.

Second, to further characterize the variability in sediment discharge capacity in subglacial compared to subaerial channels, $100$  parameter combinations are applied to a range of different channel geometry factors ($\beta$ and $k$) and friction factors ($f_r$ and $f_p$). \mauro{So how many runs are actually run?  Not clear.}
Runs with parameter combinations are culled if their mean subglacial water velocity over the season lies outwith $0.75$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $2$\,\unit{m}$^3$\,\unit{s}$^{-1}$ or if subaerial water velocity lies outwith $0.3$\,\unit{m}$^3$\,\unit{s}$^{-1}$ and $1.2$\,\unit{m}$^3$\,\unit{s}$^{-1}$, in accordance with measurements \cite<e.g.>{werder2010b,magnusson2012,chandler2013} \mauro{Awkward sentence.  (also note that ``cull'' means to remove, thus I changed ``between'' to ``outwith''.}.
The variability of output is represented by the  standard deviation of that output over the course of the season \mauro{This sentence needs context.  What is variability used for?}.

Lastly, outputs of $w\tau$ are compared with the sediment discharge records of the glaciers using Spearman rank correlation, dependent on the ordering of values.
Rank correlation reduces the impact of the non-linear response of sediment transport capacity to the hydrological forcing, thus making our interpretation independent of the exact details of the sediment transport formula.
Outputs and observations, which have a $15$\,\unit{min} resolution, are aggregated across time intervals of up to $15$\,\unit{days} by averaging values over a symmetric window of corresponding size.

@@

\subsection{Role of water discharge in sediment transport capacity}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig2.pdf}
  \caption{Parameterization outputs from the hydrographs n panels a and e for \alpine scenarios (a-d) and \icesheet{} (e-h).
    Gray (black) lines in a and e represent the subglacial channel size (water discharge).
    Blue lines represent outputs from the subglacial channel, while orange lines represent the subaerial channel.
    Data are shown at $15$\unit{min} intervals, i.e. not aggregated.
    Arrows denote where variables peaking in the subglacial channel prior to the subaerial one \mauro{wouldn't this need more zoomed-in plots?}.
    Insets in e-h show the  peak melt event \icesheet{} denoted by the shaded area in panel e.
    % Note that the x and y axes are different for \alpine{} and \icesheet{}.
        \mauro{Wrong units! Shear stress is Pa\,m$^{-2}$ and width integrated Pa\,m$^{-1}$}
  }
  \label{fig:model_outs}
\end{figure}

\mauro{Good para}
Following the parameterizations defined in Sections~\ref{sect:sub_mode}~and~\ref{sect:fluv}, the water velocity, shear stress, and width-integrated shear stress exhibit different seasonal evolutions and peaks.
Water velocity and shear stress represent the potential for sediment mobilization, and width-integrated shear stress represents a proxy for the total sediment transport capacity across the channel bed (Figure~\ref{fig:model_outs}).

\mauro{In principle good paragraph, the meat.  But suffers from sloppy writing.}
In the subglacial channel, peaks in $w_g\tau_g$ generally occur when water discharge increases at the fastest rate (Figure~\ref{fig:model_outs}d,h).
In contrast, in the subaerial channel,   $w_g\tau_g$  peaks when water discharge is highest.
The timing of the greatest rate of change in water discharge occurs before the peaks in water discharge \mauro{what?}.
Furthermore, the subglacial outputs \mauro{What are ``outputs''?}  are generally larger than the subaerial values, thus more efficient at transporting sediment.
Here, sediment could be deposited in the proglacial area as the water depressurizes, then remobilized when sediment transport capacity peaks in the subaerial system \mauro{not sure this tangent is worth taking in this paragraph}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.9\linewidth]{Fig3.png}
  \caption{
    Relationship between water discharge and velocity, shear stress, and width-integrated shear stress for \alpine{} (a-c) and \icesheet (d-f).
    These outputs' relationship with measured sediment discharge is given for \alpine in (g-i) and \icesheet (j-l).
    Variables on the x-axis have been normalized to mean values.
    $R_g$ and $R_f$ show the Spearman rank correlations for the subglacial and the subaerial outputs, respectively.
    Plots are shown with data and outputs at $15$\unit{min} intervals.
    \mauro{Pixely, needs more resolution, at least 2x. The two dots in the legend are not visible.  I would add ``Measured'' to the two y-labels.}
  }
  \label{fig:Qw_vari}
\end{figure}
